---
title: "script"
author: "Charles"
date: "16 8 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Description

In this analysis of the [GSE141259](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE141259) dataset, I want to clarify the identity of the M2 and resolving macrophages, which the author does not specify.

## Setup

```{r setup, include=FALSE}
.libPaths("V:/GROUPS/AG Ramming/Fellows/Charles/R-4.1.1/library")
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

## Load the libraries

```{r libraries}
library(SeuratObject)
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(openxlsx)
library(SeuratData)
library(SeuratDisk)
library(randomcoloR)
```

## load the dataset and perform differential expression analysis.

```{r gse141259}
set.seed(123)
colours <- distinctColorPalette(200)

obj_path <- "V:/GROUPS/AG Ramming/Fellows/Charles/Analyses/P01-ATF3 in ECs-Charles/GSE141259-mouse/QC_clustering_annotation_DEA/objects/analyses/version1"

object_name <- "integrated_and_clustered.rds"

seurat_obj <- readRDS(paste0(obj_path, "/", object_name))

# inspect the seurat object
DimPlot(seurat_obj, group.by = "metacelltype")

unique(seurat_obj$metacelltype)

table(seurat_obj$cell.type, seurat_obj$group)

# identify markers of the metacelltypes
Idents(seurat_obj) <- seurat_obj$metacelltype

cell_markers <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data")
write.xlsx(cell_markers,
           "results/metacell_markers.xlsx",
           rowNames = TRUE)

cell_markers_sig <- subset(cell_markers,
                           p_val_adj < 0.05 & avg_log2FC > 0)

top_cell_markers <- cell_markers_sig %>%
  group_by(cluster) %>% top_n(5, avg_log2FC)

# visualise with heatmap
DefaultAssay(seurat_obj) <- "RNA"

Idents(seurat_obj) <- seurat_obj$metacelltype

seurat_obj <- ScaleData(seurat_obj, assay = "RNA",
                        features = rownames(seurat_obj))

Seurat::DoHeatmap(seurat_obj, assay = "RNA",
          features = top_cell_markers$gene, label = FALSE,
          group.colors = colours) +
  scale_color_manual(values = colours)
ggsave("figures/heatmap_metacell_markers.pdf",
       height = 8, width = 8, dpi = 500)

# identify markers of the sub celltypes
Idents(seurat_obj) <- seurat_obj$cell.type

sub_cell_markers <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data")
write.xlsx(sub_cell_markers,
           "results/sub_cell_type_markers.xlsx",
           rowNames = TRUE)

# identify markers of the sub celltypes version2
# this includes ILC2 annotation

Idents(seurat_obj) <- seurat_obj$cell.type_2

sub_cell_markers2 <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data",
                               only.pos = TRUE)
write.xlsx(sub_cell_markers2,
           "results/sub_cell_type_markers_with_ILC2.xlsx",
           rowNames = TRUE)
```

Results from the gene enrichment analysis show that the author's identification of the cells is consistent with literature and reliable.

But because the the use ofthe terminology, M1 and M2 macrophages are obsolete, I will recluster the macrophages to better identify the macrophage subtypes.

## subset the macrophages and recluster them.

```{r macrophages}
# subset the macrophages
Idents(seurat_obj) <- seurat_obj$metacelltype

macrophages <- subset(seurat_obj,
                      idents = "macrophages")

# I consider the day to be the main source of batch

# Split seurat object by condition to perform cell cycle scoring and SCT on all samples
split_seurat <- SplitObject(macrophages, 
                            split.by = "day")

# use cconvert_human_to_mouse_symbols from nichenetr
library(nichenetr)

m.g2m.genes <- convert_human_to_mouse_symbols(cc.genes.updated.2019$g2m.genes)
m.s.genes <- convert_human_to_mouse_symbols(cc.genes.updated.2019$s.genes)


#to increase limits for allowable objects

options(future.globals.maxSize = 4000 * 1024^2)

for (i in 1:length(split_seurat)) {
    split_seurat[[i]] <- NormalizeData(split_seurat[[i]], verbose = TRUE)
    split_seurat[[i]] <- CellCycleScoring(split_seurat[[i]], 
                                    g2m.features=m.g2m.genes,
                                    s.features=m.s.genes)
    
    split_seurat[[i]]$CC.Difference <- split_seurat[[i]]$S.Score - split_seurat[[i]]$G2M.Score

    split_seurat[[i]] <- SCTransform(split_seurat[[i]], 
                                     vars.to.regress = c("CC.Difference"))
}


# perform integration
# Select the most variable features to use for integration
integ_features <- SelectIntegrationFeatures(object.list = 
                                            split_seurat, 
                                            nfeatures = 3000,
                                            fvf.nfeatures = 
                                            3000) 

# Prepare the SCT list object for integration
split_seurat <- PrepSCTIntegration(object.list = split_seurat,
                                   
                                   anchor.features = 
                                     integ_features)


# data set is large so I run this before
split_seurat <- lapply(X = split_seurat, FUN = RunPCA, 
                       verbose = FALSE)

# note that I have to set the reduction parameter of the FindIntegrationAnchors function
# t to "rpca"

# Find best buddies - can take a while to run
integ_anchors <- FindIntegrationAnchors(object.list = 
                                          split_seurat, 
                                        normalization.method = "SCT", 
                                        anchor.features = integ_features,
                                        dims = 1:50,
                                        reference = c(3, 5),
                                        reduction = "rpca")

rm(merged_seurat_object, merged_seurat_object_phase)
gc()

rm(split_seurat)
# Integrate across strains
seurat_integrated <- IntegrateData(anchorset = integ_anchors, 
                                   normalization.method = "SCT",
                                   dims = 1:50)

seurat_integrated <- RunPCA(seurat_integrated)
seurat_integrated <- RunUMAP(seurat_integrated,
                       dims = 1:30, 
                       reduction = "pca")

seurat_integrated <- FindNeighbors(seurat_integrated,
                             reduction = "umap", dims = 1:2)
seurat_integrated <- FindClusters(seurat_integrated,
                            resolution = seq(0.05, 3, 0.05))
DimPlot(seurat_integrated)

DefaultAssay(seurat_integrated) <- "RNA"

seurat_integrated <- NormalizeData(seurat_integrated)

FeaturePlot(seurat_integrated,
            features = "Gpnmb", order = TRUE)

saveRDS(seurat_integrated,
        "objects/macrophages.rds")
```

## Cell clustering with IKAP algorithm

To discover major macrophages sub-pupulations, I perform clustering with IKAP.

```{r ikap_with_alra}
library(IKAP)
library(openxlsx)
library(ggplot2)

DefaultAssay(seurat_integrated) <- "integrated"

seurat_integrated <- ikap3(seurat_integrated,
                           out.dir = "./IKAP")

DefaultAssay(seurat_integrated) <- "RNA"

saveRDS(seurat_integrated,
        "objects/macrophages.rds")

# best clusterfing from IKAP is if one uses 5 PCs and 5 clusters

DefaultAssay(seurat_integrated) <- "RNA"

Idents(seurat_integrated) <- factor(seurat_integrated$PC7K8)

DimPlot(seurat_integrated, label = TRUE,
        split.by = "group")

PC7K8_markers <- FindAllMarkers(seurat_integrated,
                                assay = "RNA",
                                slot = "data",
                                only.pos = TRUE)
write.xlsx(PC7K8_markers,
           "results/PC7K8_markers.xlsx",
           rowNames = TRUE)

sig_PC7K8_markers <- subset(PC7K8_markers,
                            p_val_adj < 0.05)

top_sig_PC7K8_markers <- sig_PC7K8_markers %>%
  group_by(cluster) %>% top_n(5, avg_log2FC)

as.data.frame(top_sig_PC7K8_markers)

seurat_integrated <- ScaleData(seurat_integrated,
                        features = rownames(seurat_integrated))

DimPlot(seurat_integrated, group.by = "PC7K8",
        cols = colours, split.by = "group", label = TRUE)

DoHeatmap(seurat_integrated, 
          features = top_sig_PC7K8_markers$gene,
          group.colors = colours)+
  scale_color_manual(values = colours)

FeaturePlot(seurat_integrated, features = "Mmp12",
            order = TRUE, split.by = "group")
```

## Comparing author's annotation of macrophages with mine

### Author's justification of M2 macrophages

### Identification of markers.

```{r author_annotation}
set.seed(123)
colours <- distinctColorPalette(200)

seurat_integrated <- readRDS("objects/macrophages.rds")

seurat_integrated$group <- factor(seurat_integrated$group,
                                  levels = c("PBS", "bleo"))

DimPlot(seurat_integrated, group.by = "cell.type",
        cols = colours) +
  theme(aspect.ratio = 1)
ggsave("figures/umap_author_MF_subtypes.pdf",
       height = 5, width = 5, dpi = 500)

DimPlot(seurat_integrated, group.by = "cell.type",
        cols = colours, split.by = "group") +
  theme(aspect.ratio = 1)
ggsave("figures/umap_author_MF_subtypes2.pdf",
       height = 6.5, width = 6.5, dpi = 500)

# get markers based on author's mf annotation

Idents(seurat_integrated) <- seurat_integrated$cell.type

auth_celltype_markers <- FindAllMarkers(seurat_integrated,
                                        assay = "RNA",
                                        slot = "data",
                                        only.pos = TRUE,
                                        min.pct = 0,
                                        logfc.threshold = 0)
write.xlsx(auth_celltype_markers,
           "results/auth_MF_subtype_markers.xlsx",
           rowNames = TRUE)

auth_celltype_markers_sig <- subset(auth_celltype_markers,
                                    p_val_adj < 0.05)

# plot mf markers
top_auth_mf <- auth_celltype_markers_sig %>%
  group_by(cluster) %>% top_n(6, avg_log2FC)

seurat_integrated <- ScaleData(seurat_integrated,
                               features = 
                               rownames(seurat_integrated),
                               assay = "RNA")

DoHeatmap(seurat_integrated, 
          features = top_auth_mf$gene,
          label = FALSE, group.colors = colours) +
    scale_color_manual(values = colours)
ggsave("figures/heatmap_author_MF_subtypes2.pdf",
       height = 7, width = 7, dpi = 500)
```

### Obtain pathways specifically enriched in the MF subtypes

```{r specific}
list.files("scripts")

source("scripts/select_specific_markers.R")

specific_markers <- select_spec_markers(auth_celltype_markers)

saveRDS(specific_markers,
        "objects/cell_spec_MF_markers_from_auth.rds")
```

### Use ORA to assess pathways enriched in M2 macrophages vs all other macrophages

```{r ORA}
library(enrichR)
dbs <- listEnrichrDbs()
tail(dbs, 100)
dbs_keep <- c("MSigDB_Hallmark_2020")

source("scripts/cell_specific_ora.R")

cell_specific_ora(marker_list =  specific_markers, 
                  dbs = dbs_keep)
```

Some paper have associated Arg1-expressing tumour associated macrophages that help in tumour resolution. Studies also show that M2 MFs are pro-resolving. After performing over-representation analysis on genes specifically enriched in M2 MFs compared to the other MFs,  results in this analysis are consistent with the fact the M2 MFs in fibrosis are foes rather than friends for the following reasons.

- They are the only MFs secreting factors that promote EMT (epithelial to mesenchymal transition).
- They are amplified during lung fibrosis.
- They are enriched with oxidative phosphorylation, which accoridng to this [paper](https://jpet.aspetjournals.org/content/379/3/290.long) is a known and targetable pathogenic pathway in lung fibrosis.
- The metabolic shift which drives of fibroblasts and relevant cells to convert to myofibroblasts requires enhanced aerobic glycolysis, lipogenesis, OXPHOS, and anaplerosis (e.g., glutaminolysis) and is evident in lung regions of active fibrosis in patients with IPF.
- M2 macrophages co-appear with myofibroblasts at day 10.

## Using ORA to identify all pathways enriched in res. Mfs.

```{r res_mfpways}
library(enrichR)

dbs <- listEnrichrDbs()

tail(dbs, 100)

dbs_keep <- c("MSigDB_Hallmark_2020",
              "GO_Biological_Process_2021",
              "GO_Cellular_Component_2021",
              "GO_Molecular_Function_2021",
              "KEGG_2019_Mouse")

list.files("results")

markers <- read.xlsx("results/auth_MF_subtype_markers.xlsx",
                     rowNames = TRUE)

markers_res_mf <- subset(markers,
                         cluster == "Resolution macrophages")

all(markers_res_mf$avg_log2FC > 0) # T

res_mf_marker_genes <- markers_res_mf$gene

res_mf_pways <- enrichr(res_mf_marker_genes, dbs_keep)

write.xlsx(res_mf_pways,
           "results/pways_res_mfs.xlsx",
           rowNames = TRUE)

for (pway in names(res_mf_pways)) {
  
  pway_oi <- res_mf_pways[[pway]]
  
  pway_oi_sig <- subset(pway_oi,
                        Adjusted.P.value < 0.05)
  
  if(!is.null(pway_oi_sig) & nrow(pway_oi_sig) != 0){
    
    plotEnrich(pway_oi) +
      ggtitle(paste0("res_mf_", pway)) +
      scale_x_discrete(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0))
    ggsave(paste0("figures/res_mf_enriched_",
                  pway, ".pdf"))
    
    
  }

}

list.files("objects")

mf_specific_mkers <- readRDS("objects/cell_spec_MF_markers_from_auth.rds")

res_mf_spec_mk_genes <- mf_specific_mkers$`Resolution macrophages`

## group specific markers using GO_CC

res_mf_cc <- enrichr(res_mf_spec_mk_genes, "GO_Cellular_Component_2021")

write.xlsx(res_mf_cc,
           "results/res_mf_cc.xlsx",
           rowNames = TRUE)

plotEnrich(res_mf_cc)

# make heatmap of res_mf_spec_mk_genes

res_mf_pways_keep <- subset(markers_res_mf,
                            gene %in% res_mf_spec_mk_genes)

res_mf_pways_keep$ratio <- res_mf_pways_keep$pct.1 /
  res_mf_pways_keep$pct.2

res_mf_pways_keep <- res_mf_pways_keep[order(res_mf_pways_keep$ratio,
                                             decreasing = T), ]

top_res_mf <- res_mf_pways_keep %>% top_n(50,
                                          avg_log2FC)

seurat_integrated <- readRDS("objects/macrophages.rds")

seurat_integrated <- ScaleData(seurat_integrated,
                               features = rownames(seurat_integrated))

DoHeatmap(seurat_integrated,
          features = top_res_mf$gene,
          label = FALSE,
          group.colors = colours) +
  scale_color_manual(values = colours)

VlnPlot(seurat_integrated,
        features = c("Folr1", "Procr",
                     "H2-M2", "Mmp12")) +
  scale_color_manual(values = colours)

```

### Use ORA to assess pathways upregulated in M2 macrophages bleo vs PBS

```{r upreg_ora}
set.seed(123)
colours <- distinctColorPalette(200)

seurat_integrated <- readRDS("objects/macrophages.rds")

seurat_integrated$group <- factor(seurat_integrated$group,
                                  levels = c("PBS", "bleo"))

Idents(seurat_integrated) <- seurat_integrated$cell.type

M2 <- subset(seurat_integrated,
            idents = "M2 macrophages")

Idents(M2) <- "group"

M2_bleo_vs_healthy_degs <- FindMarkers(M2,
                                          assay = "RNA",
                                          slot = "data",
                                          ident.1 = "bleo")
write.xlsx(M2_bleo_vs_healthy_degs,
           "results/M2_bleo_vs_PBS_degs.xlsx",
           rowNames = TRUE)

M2_sig <- subset(M2_bleo_vs_healthy_degs,
                 p_val_adj < 0.05 & avg_log2FC > 0)

M2_sig # very few DEGs.

table(seurat_integrated$cell.type, seurat_integrated$group)
```

There are few DEGs in M2 when we compare them in bleo vs in the PBS conditions because there are very few M2s (56) in the healthy condition compared to 2306 in the fibrotic condition.

### Use ORA to assess pathways upregulated in resolution macrophages

```{r resol_mfs}
list.files("objects")

set.seed(123)
colours <- distinctColorPalette(200)

seurat_integrated <- readRDS("objects/macrophages.rds")

seurat_integrated$group <- factor(seurat_integrated$group,
                                  levels = c("PBS", "bleo"))

Idents(seurat_integrated) <- seurat_integrated$cell.type

res_mf <- subset(seurat_integrated,
            idents = "Resolution macrophages")

```

## My clustering

I determined my best clusters using the IKAP algorithm. This algorithm showed that 7 PCs and 8 clusters best represented cell populations within the macrophage sub-population.

### Find out markers for each of the IKAP clusters.

```{r my_clustering}
set.seed(123)
colours <- distinctColorPalette(200)

seurat_integrated <- readRDS("objects/macrophages.rds")

seurat_integrated$group <- factor(seurat_integrated$group,
                                  levels = c("PBS", "bleo"))

Idents(seurat_integrated) <- factor(as.numeric(seurat_integrated$PC19K11))

pc19k11_markers <- FindAllMarkers(seurat_integrated,
                                  assay = "RNA",
                                  slot = "data",
                                  only.pos = TRUE)

clust4 <- subset(pc19k11_markers,
                 cluster==4)

DimPlot(seurat_integrated, 
        cols = colours, label = TRUE, split.by = "group") +
  theme(aspect.ratio = 1)
ggsave("figures/umap_author_MF_subtypes.pdf",
       height = 5, width = 5, dpi = 500)

PC7K8_markers <- read.xlsx("results/PC7K8_markers.xlsx",
                            rowNames = TRUE)

sig_PC7K8_markers <- subset(PC7K8_markers,
                            p_val_adj < 0.05)

top_sig_PC7K8_markers <- sig_PC7K8_markers %>%
  group_by(cluster) %>% top_n(7, avg_log2FC)

as.data.frame(top_sig_PC7K8_markers)

DoHeatmap(seurat_integrated, 
          features = top_sig_PC7K8_markers$gene,
          group.colors = colours)+
  scale_color_manual(values = colours)
```

After my own exploration, I decide to stick with the annotation provided by the authors.

## Comparing enriched pathways in the various clusters

### First convert gene symbole to the entrez gene names

```{r entrez}
library(org.Mm.eg.db)

res_tableOE <- read.xlsx("results/auth_MF_subtype_markers.xlsx",
                           rowNames = TRUE)

## clusterProfiler does not work as easily using gene names, so we will turn gene names into Ensembl IDs using 
## clusterProfiler::bitr and merge the IDs back with the DE results

ids <- bitr(res_tableOE$gene, 
            fromType = "SYMBOL", 
            toType = c("ENSEMBL", "ENTREZID"), 
            OrgDb = "org.Mm.eg.db")
# 0.48% of input gene IDs are fail to map...

## The gene names can map to more than one Ensembl ID (some genes change ID over time), 
## so we need to remove duplicate IDs prior to assessing enriched GO terms
non_duplicates <- which(duplicated(ids$SYMBOL) == FALSE)

ids <- ids[non_duplicates, ] 

## Merge the Ensembl IDs with the results     
merged_gene_ids <- merge(x=res_tableOE, y=ids, by.x="gene", by.y="SYMBOL")  

write.xlsx(merged_gene_ids,
           "results/auth_MF_subtype_markers_with_entrez.xlsx",
           rowNames = TRUE)
colnames(merged_gene_ids)
```

### Obtain entrez gene names for each of the marker genes

```{r obtaining_entrez}
list.files("scripts")
source("scripts/select_spec_entrez_markers.R")

res_tableOE <- read.xlsx("results/auth_MF_subtype_markers_with_entrez.xlsx",
                           rowNames = TRUE)

entrez_markers <- select_spec_entrez_markers(res_tableOE)

saveRDS(entrez_markers,
        "objects/auth_MF_subtype_markers_with_entrez.rds")

# all significant cell markers

list.files("scripts")
source("scripts/select_all_signif_entrez_markers.R")

entrez_markers2 <- select_all_sig_entrez_markers(res_tableOE)

saveRDS(entrez_markers2,
        "objects/auth_all_MF_subtype_markers_with_entrez.rds")
```

## Compare the various macrophage subclusters, based on their specific marker genes.

### Load the marker genes

```{r load_geneset}
list.files("objects")

library(clusterProfiler)
library(org.Mm.eg.db)

specific_entrez <- readRDS("objects/auth_all_MF_subtype_markers_with_entrez.rds")

R.utils::setOption("clusterProfiler.download.method","auto")
getOption("clusterProfiler.download.method")

ck <- compareCluster(geneCluster = specific_entrez, 
                     fun = enrichPathway, 
                     pvalueCutoff = 0.05,
                     qvalueCutoff = 0.05,
                      organism = "mmu")

ck <- setReadable(ck, 
                  OrgDb = org.Mm.eg.db, 
                  keyType="ENTREZID")
head(ck) 
# View(ck)
head(as.data.frame(ck))

dotplot(ck) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5), aspect.ratio = 1)

# with enrichGO
enGO <- compareCluster(geneCluster = specific_entrez, 
                         fun = "enrichGO", 
                         pvalueCutoff = 0.05,
                         qvalueCutoff = 0.05,
                         OrgDb = org.Mm.eg.db)

enGO <- setReadable(enGO, 
                    OrgDb = org.Mm.eg.db, 
                    keyType="ENTREZID")
head(enGO) 
# View(enGO)
head(as.data.frame(enGO))

dotplot(enGO, showCategory = 5) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5), aspect.ratio = 1)


# with enrichPathway
require(ReactomePA)

R.utils::setOption("clusterProfiler.download.method","auto")
getOption("clusterProfiler.download.method")

enReactome <- compareCluster(geneCluster = specific_entrez, 
                         fun = "enrichPathway", 
                         pvalueCutoff = 0.05,
                         qvalueCutoff = 0.05,
                         organism = "mouse",
                         readable = TRUE)
# View(enReactome)
head(as.data.frame(enReactome))

dotplot(enReactome, showCategory = 5) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5), aspect.ratio = 1)
```

### identify signalling pathways that change in resolving macrophages from day 3 to 28 of lung fibrosis.

```{r load_gsva}
library(escape)

path_to_gsva <- "V:/GROUPS/AG Ramming/Fellows/Charles/Analyses/P01-ATF3 in ECs-Charles/GSE141259-mouse/gsva/objects"

gsva_file <- "gsva_obj_df_with_day.xlsx"

gsva_df <- read.xlsx(paste0(path_to_gsva,
                            "/", gsva_file),
                     rowNames = TRUE)

head(gsva_df[1:10, 1:10])

# subset only res_mfs

resl_mfs <- subset(gsva_df,
                   celltype == "Resolution macrophages")

# only bleo
resl_mfs_bleo <- subset(resl_mfs,
                        group == "bleo")

table(resl_mfs_bleo$celltype)

# remove irrelevant columns
resl_mfs_bleo$celltype <- NULL
resl_mfs_bleo$group <- NULL
resl_mfs_bleo$replicate <- NULL

class(resl_mfs_bleo$day)

resl_mfs_bleo$day <- factor(resl_mfs_bleo$day)

write.csv(resl_mfs_bleo,
          "results/resl_mfs_bleo.csv",
          quote = FALSE,
          row.names = TRUE,
          col.names = TRUE)

output <- getSignificance(resl_mfs_bleo, 
                          group = "day", 
                          fit = "ANOVA")

output <- output[order(output$FDR), ]

View(output)

View(output)

colnames(resl_mfs_bleo)[grep("DEGRA", 
                             colnames(resl_mfs_bleo))]

output["REACTOME_COLLAGEN_DEGRADATION", ]
```

## Plotting the development of interesting pathways on resolution MFs.

```{r cd44_pways}
library(escape)

path_to_gsva <- "V:/GROUPS/AG Ramming/Fellows/Charles/Analyses/P01-ATF3 in ECs-Charles/GSE141259-mouse/gsva/objects"

gsva_file <- "gsva_obj_df_with_day.xlsx"

gsva_df <- read.xlsx(paste0(path_to_gsva,
                            "/", gsva_file),
                     rowNames = TRUE)

head(gsva_df[1:10, 1:10])


# according to reactome 
# cd44 is involved in
cd44_pways <- c("",
"REACTOME_COLLAGEN_DEGRADATION")

gsva_cd44 <- resl_mfs[,
                         c(cd44_pways, "day", "group")]

gsva_cd44 <- gsva_cd44 %>%
  gather(cd44_pways, 
         key = "pways", value = "score")
# View(gsva_cd44)

gsva_cd44_keep <- gsva_cd44 %>%
  group_by(pways, day, group) %>%
  summarise(the_median = median(score))

gsva_cd44_keep$group <- factor(gsva_cd44_keep$group,
                               levels = c("PBS", "bleo"))

ggplot(gsva_cd44_keep, 
       aes(x = day, y = the_median, 
           color = pways, group = pways)) + 
  geom_point() + 
  geom_line() + 
  facet_grid(~ group) +
  expand_limits(x=c(0,30), y=c(0.10, 1)) +
  theme_bw()


library(randomcoloR)
set.seed(123)
colours <- distinctColorPalette(300)

days <- unique(gsva_cd44$day)

for (day_oi in days) {
day_df <- subset(gsva_cd44,
               day == day_oi)

day_df$group <- factor(day_df$group,
                     levels = c("PBS", "bleo"))

ggplot(day_df, aes(celltype, score, fill = pways))+
  geom_bar(stat = "identity", show.legend = FALSE) +
  ggtitle(paste0("day_", day_oi)) +
  facet_grid(~ group) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
 scale_fill_manual(values = colours) 
ggsave(paste0("figures/day_", day_oi, ".pdf"),
       height = 5, width = 9, dpi = 500)
}

library(ggpubr)

day_df <- subset(gsva_cd44,
               day == day_oi)

day_df$group <- factor(day_df$group,
                     levels = c("PBS", "bleo"))

p <- ggplot(day_df, aes(celltype, score, fill = pways))+
  geom_bar(stat = "identity") +
  ggtitle(paste0("day_", day_oi)) +
  facet_grid(~ group) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
 scale_fill_manual(values = colours) 

# Extract the legend. Returns a gtable
leg <- get_legend(p)

# Convert to a ggplot and print
as_ggplot(leg)

ggsave("figures/legend.pdf",
       height = 5, width = 5, dpi = 300)

gsva_cd44_bleo <- subset(gsva_cd44,
                         group == "bleo")

View(gsva_cd44_bleo)

gsva_cd44_bleo <- gsva_cd44 %>% group_by(day, pways) %>%
  summarise(the_median = median(score))

library(randomcoloR)
set.seed(123)
colours <- distinctColorPalette(300)

ggplot(gsva_cd44_bleo, aes(x=day, y=the_median,
                           fill = pways, group = "pways")) +
  geom_point(aes(color = pways)) +
  geom_line(aes(color = pways)) +
  # geom_line(aes(color = pways), size = 1) +
  scale_fill_manual(values = colours) +
  # scale_color_manual(values = colours) +
  theme_minimal()
```

## Match collagen production in fibroblasts and collagen degradation by macophages.

```{r fib_resmf}
library(escape)

path_to_gsva <- "V:/GROUPS/AG Ramming/Fellows/Charles/Analyses/P01-ATF3 in ECs-Charles/GSE141259-mouse/gsva/objects"

gsva_file <- "gsva_obj_df_with_day.xlsx"

gsva_df <- read.xlsx(paste0(path_to_gsva,
                            "/", gsva_file),
                     rowNames = TRUE)

head(gsva_df[1:10, 1:10])

colnames(gsva_df)[grep("COLLAG", 
                             colnames(gsva_df))]

pway_oi <- c("REACTOME_COLLAGEN_FORMATION",
             "REACTOME_COLLAGEN_CHAIN_TRIMERIZATION",
             "REACTOME_COLLAGEN_BIOSYNTHESIS_AND_MODIFYING_ENZYMES",
             "REACTOME_ECM_PROTEOGLYCANS",
           "REACTOME_COLLAGEN_DEGRADATION",
           "REACTOME_CROSSLINKING_OF_COLLAGEN_FIBRILS")

colnames(gsva_df)[grep("ECM", 
                             colnames(gsva_df))]

unique(gsva_df$celltype)

cells_oi <- subset(gsva_df,
                   celltype %in% 
                     c("Resolution macrophages", 
                       "Myofibroblasts"))

cells_oi <- cells_oi[, c(pway_oi, "day", "group",
                         "celltype")]

to_plot <- cells_oi %>%
  gather(pway_oi, 
         key = "pways", value = "score")

to_plot_keep <- to_plot %>%
  group_by(pways, day, group) %>%
  summarise(the_median = median(score))

to_plot_keep$group <- factor(to_plot_keep$group,
                               levels = c("PBS", "bleo"))

ggplot(to_plot_keep, 
       aes(x = day, y = the_median, 
           color = pways, group = pways)) + 
  geom_point() + 
  geom_line() + 
  facet_grid(~ group) +
  expand_limits(x=c(0,30), y=c(0.10, 1)) +
  theme_bw() + NoLegend()
```

### Find out the 5 most inportant signalling pathways in resolution macrophages and how they vary from days 3 through 28 during fibrosis.

### By Variance threshold

```{python VarThresh}
import pandas as pd

X = pd.read_csv("results/resl_mfs_bleo.csv",
          sep = ",",index_col = 0)

X.head()

# remove the day column as it is categorical
X_num = X.drop(columns=['day'])
 
print(X_num.head())

# import modules
from sklearn.feature_selection import VarianceThreshold

# remove zero-variance features 
selector = VarianceThreshold(threshold=0)  # 0 is default
 
print(selector.fit_transform(X_num))

# Specify `indices=True` to get indices of selected features
print(selector.get_support(indices=True))

# Use indices to get the corresponding column names of selected features
num_cols = list(X_num.columns[selector.get_support(indices=True)])
 
print(num_cols)

# Subset `X_num` to retain only selected features
X_num = X_num[num_cols]
 
print(X_num)
```

Results from the variance analysis shows that only one feature has a variance of 0, from days 3 through 28 of fibrosis.

### Mutual information classification

```{python MIC}
import pandas as pd

X = pd.read_csv("results/resl_mfs_bleo.csv",
          sep = ",",index_col = 0)

from sklearn.preprocessing import LabelEncoder
 
le = LabelEncoder()

# remove day, the categorical variable
X_num = X.drop(columns=['day'])

print(X_num)

from sklearn.preprocessing import LabelEncoder
 
le = LabelEncoder()

y = X['day']
 
print(y)

y = le.fit_transform(y)

from sklearn.feature_selection import mutual_info_classif
 
print(mutual_info_classif(X_num, y, random_state=68))

from sklearn.feature_selection import SelectKBest
from functools import partial
 
score_func = partial(mutual_info_classif, random_state=68)
 
# Select top 10 features with the most mutual information
selection = SelectKBest(score_func=score_func, k=10)
 
print(selection.fit_transform(X_num, y))

X = X_num[X_num.columns[selection.get_support(indices=True)]]
 
print(X.columns)
```

### Mutual information regression

```{python MIC}
import pandas as pd

X = pd.read_csv("results/resl_mfs_bleo.csv",
          sep = ",",index_col = 0)

from sklearn.preprocessing import LabelEncoder
 
le = LabelEncoder()

# remove day, the categorical variable
X_num = X.drop(columns=['day'])

print(X_num)

from sklearn.preprocessing import LabelEncoder
 
le = LabelEncoder()

y = X['day']
 
print(y)

from sklearn.feature_selection import mutual_info_regression
 
print(mutual_info_regression(X_num, y, random_state=68))

from sklearn.feature_selection import SelectKBest
from functools import partial
 
score_func = partial(mutual_info_regression, random_state=68)
 
# Select top 10 features with the most mutual information
selection = SelectKBest(score_func=score_func, k=10)
 
print(selection.fit_transform(X_num, y))

X = X_num[X_num.columns[selection.get_support(indices=True)]]
 
print(X.columns)
```


### By SFFS

```{python SFFS}
import pandas as pd
from sklearn.linear_model import LogisticRegression
from mlxtend.feature_selection import SequentialFeatureSelector as SFS
import pandas as pd
import matplotlib.pyplot as plt
from mlxtend.plotting import plot_sequential_feature_selection as plot_sfs

X = pd.read_csv("results/resl_mfs_bleo.csv",
          sep = ",",index_col = 0)

# remove day, the categorical variable
X_num = X.drop(columns=['day'])

y = X['day']

lr = LogisticRegression(max_iter=1000,
multi_class='multinomial', solver='lbfgs')

# Sequential forward selection
sfs = SFS(lr,
          k_features=5,
          forward=True,
          floating=True,
          scoring='accuracy',
          cv=0)
sfs.fit(X_num, y)
# Print the chosen feature names
print(sfs.subsets_)
# Print the accuracy of the model after sequential forward selection
print(sfs.subsets_[3]['avg_score'])

# Plot the model accuracy
plot_sfs(sfs.get_metric_dict())
plt.show()

```

## Session infomation
```{r session_info}
sessionInfo()
```
