---
title: "script"
author: "Charles"
date: "16 8 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Description

In this analysis of the [GSE141259](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE141259) dataset, I want to clarify the identity of the M2 and resolving macrophages, which the author does not specify.

## Setup

```{r setup, include=FALSE}
.libPaths("V:/GROUPS/AG Ramming/Fellows/Charles/R-4.1.1/library")
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

## Load the libraries

```{r libraries}
library(SeuratObject)
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(openxlsx)
library(SeuratData)
library(SeuratDisk)
library(randomcoloR)
```

## load the dataset and perform differential expression analysis.

```{r gse141259}
set.seed(123)
colours <- distinctColorPalette(200)

obj_path <- "V:/GROUPS/AG Ramming/Fellows/Charles/Analyses/P01-ATF3 in ECs-Charles/GSE141259-mouse/QC_clustering_annotation_DEA/objects/analyses/version1"

object_name <- "integrated_and_clustered.rds"

seurat_obj <- readRDS(paste0(obj_path, "/", object_name))

# inspect the seurat object
DimPlot(seurat_obj, group.by = "metacelltype")

unique(seurat_obj$metacelltype)

table(seurat_obj$cell.type, seurat_obj$group)

# identify markers of the metacelltypes
Idents(seurat_obj) <- seurat_obj$metacelltype

cell_markers <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data")
write.xlsx(cell_markers,
           "results/metacell_markers.xlsx",
           rowNames = TRUE)

cell_markers_sig <- subset(cell_markers,
                           p_val_adj < 0.05 & avg_log2FC > 0)

top_cell_markers <- cell_markers_sig %>%
  group_by(cluster) %>% top_n(5, avg_log2FC)

# visualise with heatmap
DefaultAssay(seurat_obj) <- "RNA"

Idents(seurat_obj) <- seurat_obj$metacelltype

seurat_obj <- ScaleData(seurat_obj, assay = "RNA",
                        features = rownames(seurat_obj))

Seurat::DoHeatmap(seurat_obj, assay = "RNA",
          features = top_cell_markers$gene, label = FALSE,
          group.colors = colours) +
  scale_color_manual(values = colours)
ggsave("figures/heatmap_metacell_markers.pdf",
       height = 8, width = 8, dpi = 500)

# identify markers of the sub celltypes
Idents(seurat_obj) <- seurat_obj$cell.type

sub_cell_markers <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data")
write.xlsx(sub_cell_markers,
           "results/sub_cell_type_markers.xlsx",
           rowNames = TRUE)

# identify markers of the sub celltypes version2
# this includes ILC2 annotation

Idents(seurat_obj) <- seurat_obj$cell.type_2

sub_cell_markers2 <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data",
                               only.pos = TRUE)
write.xlsx(sub_cell_markers2,
           "results/sub_cell_type_markers_with_ILC2.xlsx",
           rowNames = TRUE)
```

Results from the gene enrichment analysis show that the author's identification of the cells is consistent with literature and reliable.

But because the the use ofthe terminology, M1 and M2 macrophages are obsolete, I will recluster the macrophages to better identify the macrophage subtypes.

## Session infomation
```{r session_info}
sessionInfo()
```
