---
title: "script"
author: "Charles"
date: "16 8 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Description

In this analysis of the [GSE141259](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE141259) dataset, I want to clarify the identity of the M2 and resolving macrophages, which the author does not specify.

## Setup

```{r setup, include=FALSE}
.libPaths("V:/GROUPS/AG Ramming/Fellows/Charles/R-4.1.1/library")
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)
```

## Load the libraries

```{r libraries}
library(SeuratObject)
library(Seurat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(openxlsx)
library(SeuratData)
library(SeuratDisk)
library(randomcoloR)
```

## load the dataset and perform differential expression analysis.

```{r gse141259}
set.seed(123)
colours <- distinctColorPalette(200)

obj_path <- "V:/GROUPS/AG Ramming/Fellows/Charles/Analyses/P01-ATF3 in ECs-Charles/GSE141259-mouse/QC_clustering_annotation_DEA/objects/analyses/version1"

object_name <- "integrated_and_clustered.rds"

seurat_obj <- readRDS(paste0(obj_path, "/", object_name))

# inspect the seurat object
DimPlot(seurat_obj, group.by = "metacelltype")

unique(seurat_obj$metacelltype)

table(seurat_obj$cell.type, seurat_obj$group)

# identify markers of the metacelltypes
Idents(seurat_obj) <- seurat_obj$metacelltype

cell_markers <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data")
write.xlsx(cell_markers,
           "results/metacell_markers.xlsx",
           rowNames = TRUE)

cell_markers_sig <- subset(cell_markers,
                           p_val_adj < 0.05 & avg_log2FC > 0)

top_cell_markers <- cell_markers_sig %>%
  group_by(cluster) %>% top_n(5, avg_log2FC)

# visualise with heatmap
DefaultAssay(seurat_obj) <- "RNA"

Idents(seurat_obj) <- seurat_obj$metacelltype

seurat_obj <- ScaleData(seurat_obj, assay = "RNA",
                        features = rownames(seurat_obj))

Seurat::DoHeatmap(seurat_obj, assay = "RNA",
          features = top_cell_markers$gene, label = FALSE,
          group.colors = colours) +
  scale_color_manual(values = colours)
ggsave("figures/heatmap_metacell_markers.pdf",
       height = 8, width = 8, dpi = 500)

# identify markers of the sub celltypes
Idents(seurat_obj) <- seurat_obj$cell.type

sub_cell_markers <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data")
write.xlsx(sub_cell_markers,
           "results/sub_cell_type_markers.xlsx",
           rowNames = TRUE)

# identify markers of the sub celltypes version2
# this includes ILC2 annotation

Idents(seurat_obj) <- seurat_obj$cell.type_2

sub_cell_markers2 <- FindAllMarkers(seurat_obj,
                               assay = "RNA",
                               slot = "data",
                               only.pos = TRUE)
write.xlsx(sub_cell_markers2,
           "results/sub_cell_type_markers_with_ILC2.xlsx",
           rowNames = TRUE)
```

Results from the gene enrichment analysis show that the author's identification of the cells is consistent with literature and reliable.

But because the the use ofthe terminology, M1 and M2 macrophages are obsolete, I will recluster the macrophages to better identify the macrophage subtypes.

## subset the macrophages and recluster them.

```{r macrophages}
# subset the macrophages
Idents(seurat_obj) <- seurat_obj$metacelltype

macrophages <- subset(seurat_obj,
                      idents = "macrophages")

# I consider the day to be the main source of batch

# Split seurat object by condition to perform cell cycle scoring and SCT on all samples
split_seurat <- SplitObject(macrophages, 
                            split.by = "day")

# use cconvert_human_to_mouse_symbols from nichenetr
library(nichenetr)

m.g2m.genes <- convert_human_to_mouse_symbols(cc.genes.updated.2019$g2m.genes)
m.s.genes <- convert_human_to_mouse_symbols(cc.genes.updated.2019$s.genes)


#to increase limits for allowable objects

options(future.globals.maxSize = 4000 * 1024^2)

for (i in 1:length(split_seurat)) {
    split_seurat[[i]] <- NormalizeData(split_seurat[[i]], verbose = TRUE)
    split_seurat[[i]] <- CellCycleScoring(split_seurat[[i]], 
                                    g2m.features=m.g2m.genes,
                                    s.features=m.s.genes)
    
    split_seurat[[i]]$CC.Difference <- split_seurat[[i]]$S.Score - split_seurat[[i]]$G2M.Score

    split_seurat[[i]] <- SCTransform(split_seurat[[i]], 
                                     vars.to.regress = c("CC.Difference"))
}


# perform integration
# Select the most variable features to use for integration
integ_features <- SelectIntegrationFeatures(object.list = 
                                            split_seurat, 
                                            nfeatures = 3000,
                                            fvf.nfeatures = 
                                            3000) 

# Prepare the SCT list object for integration
split_seurat <- PrepSCTIntegration(object.list = split_seurat,
                                   
                                   anchor.features = 
                                     integ_features)


# data set is large so I run this before
split_seurat <- lapply(X = split_seurat, FUN = RunPCA, 
                       verbose = FALSE)

# note that I have to set the reduction parameter of the FindIntegrationAnchors function
# t to "rpca"

# Find best buddies - can take a while to run
integ_anchors <- FindIntegrationAnchors(object.list = 
                                          split_seurat, 
                                        normalization.method = "SCT", 
                                        anchor.features = integ_features,
                                        dims = 1:50,
                                        reference = c(3, 5),
                                        reduction = "rpca")

rm(merged_seurat_object, merged_seurat_object_phase)
gc()

rm(split_seurat)
# Integrate across strains
seurat_integrated <- IntegrateData(anchorset = integ_anchors, 
                                   normalization.method = "SCT",
                                   dims = 1:50)

seurat_integrated <- RunPCA(seurat_integrated)
seurat_integrated <- RunUMAP(seurat_integrated,
                       dims = 1:30, 
                       reduction = "pca")

seurat_integrated <- FindNeighbors(seurat_integrated,
                             reduction = "umap", dims = 1:2)
seurat_integrated <- FindClusters(seurat_integrated,
                            resolution = seq(0.05, 3, 0.05))
DimPlot(seurat_integrated)

DefaultAssay(seurat_integrated) <- "RNA"

seurat_integrated <- NormalizeData(seurat_integrated)

FeaturePlot(seurat_integrated,
            features = "Gpnmb", order = TRUE)
```

## Cell clustering with IKAP algorithm

## Session infomation
```{r session_info}
sessionInfo()
```
