---
title: "cellchat"
author: "Charles"
date: "11 5 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load in the libraries

```{r libraries}
.libPaths("V:/GROUPS/AG Ramming/Fellows/Charles/R-4.1.1/library")
library(Matrix)
library(SeuratObject)
library(Seurat)
library(SeuratWrappers)
library(SeuratData)
library(stringr)
library(ggplot2)
```

# load in the data set
```{r dataset}
list.files("objects")

seurat <- readRDS("objects/integrated_and_clustered_GSE141259_and_lab.rds")

DimPlot(seurat, split.by = "GSE", label = T)

DimPlot(seurat,
        group.by = "integrated_snn_res.0.2", 
        split.by = "GSE",
        label = T)
table(seurat$GSE)

cell_clusters <- table(seurat$cell.type, 
      seurat$integrated_snn_res.0.2)

cell_clusters <- t(cell_clusters)

cell_clusters <- as.data.frame.matrix(cell_clusters)

for (i in 1:nrow(cell_clusters)) {
  
  cell_clusters[i, "annot"] <- colnames(cell_clusters)[which.max(cell_clusters[i, ])]
  
}

cell_clusters

# adjust cluster 17 (row index 18) to be ILC2
cell_clusters[18, "annot"] <- "ILC2"

seurat$integrated_snn_res.0.2_modif1 <-
  as.numeric(seurat$integrated_snn_res.0.2)-1

for (i in rownames(cell_clusters)) {
  row_as_numeric <- as.numeric(i)
  seurat$integrated_snn_res.0.2_modif1[
    seurat$integrated_snn_res.0.2_modif1==i
  ] <- cell_clusters[i, "annot"]
  
}

DimPlot(seurat,
        group.by = "integrated_snn_res.0.2_modif1")

table(seurat$integrated_snn_res.0.2_modif1,
      seurat$GSE)

memory.limit(799165999999)
gc()
# perform imputation
seurat <- RunALRA(seurat)

saveRDS(seurat,
        "objects/integrated_and_clustered_GSE141259_and_lab.rds")

# write out metadata

write.xlsx(seurat@meta.data,
           "metadata/metadata.xlsx",
           rowNames=T)
```

# subset days 10 and 28 from the public and mouse data set
```{r day10 and 28}
DimPlot(seurat,
        group.by = "day")

Idents(seurat) <- "day"

days10_and_28 <- subset(seurat,
                        idents = c(10, 28))

DimPlot(days10_and_28, group.by = "day", split.by = "day")

Idents(days10_and_28) <- c("day", "GSE")

days10_and_28_keep <- days10_and_28[,
                                    days10_and_28@meta.data$day==10 | days10_and_28@meta.data$GSE=="lab"]

DimPlot(days10_and_28_keep, split.by = "day")

saveRDS(days10_and_28_keep,
        "objects/public_day10_and_lab_day28.rds")
```

# Prepare for cellchat by only subsetting those cells with a count of at least 20 in any of both conditions (bleo vs saline)

```{r subsetting}

Idents(days10_and_28_keep) <- days10_and_28_keep$group

days10_and_28_keep$group1 <- days10_and_28_keep$group

days10_and_28_keep$group1[
  days10_and_28_keep$group1=="PBS"
] <- "healthy"

days10_and_28_keep$group1[
  days10_and_28_keep$group1=="saline"
] <- "healthy"


days10_and_28_keep$group1[!days10_and_28_keep$group1=="healthy"
] <- "fibrotic"

Idents(days10_and_28_keep) <- "group1"
DimPlot(days10_and_28_keep,
        group.by = "group1")

cells_per_group <- table(days10_and_28_keep$integrated_snn_res.0.2_modif1,
                           days10_and_28_keep$group1)

days10_and_28_keep <- SCTransform(days10_and_28_keep,
                                  vars.to.regress = "GSE")
days10_and_28_keep <- RunALRA(days10_and_28_keep,
                              assay = "SCT",
                              slot = "data")
```

# Use all cells for cellchat: subset for bleo condition
```{r cellchat}
Idents(days10_and_28_keep) <- "group1"
subset_bleo <- subset(days10_and_28_keep,
                      idents = "fibrotic")

DimPlot(subset_bleo)

saveRDS(subset_bleo,
        "objects/seu_for_cellchat_fibrotic.rds")
```

# Perform cellchat
## load in the libraries
```{r cellchat}
library(CellChat)
library(patchwork)
library(NMF)
library(ComplexHeatmap)
library(ggalluvial)
options(stringsAsFactors = FALSE)
```

## Part I: Data input & processing and initialization of CellChat object
```{r input}

Idents(subset_bleo) <- subset_bleo$integrated_snn_res.0.2_modif1

DimPlot(subset_bleo, group.by = "group1")

data.input <- GetAssayData(subset_bleo, assay = "alra", slot = "data") # normalized data matrix
labels <- Idents(subset_bleo)
meta <- data.frame(labels = labels, row.names = names(labels)) # create a dataframe of the cell labels
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

cellchat <- addMeta(cellchat, meta = meta, meta.name = "labels")
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels

# set LR-interaction db
CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)


# use a subset of CellChatDB for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB, 
                           search = "Cell-Cell Contact") # use Secreted Signaling
# use all CellChatDB for cell-cell communication analysis
# CellChatDB.use <- CellChatDB # simply use the default CellChatDB
# set the used database in the object
cellchat@DB <- CellChatDB.use


# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multiprocess", workers = 4) # do parallel


cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
# project gene expression data onto PPI (Optional: when running it, USER should set `raw.use = FALSE` in the function `computeCommunProb()` in order to use the projected data)
# cellchat <- projectData(cellchat, PPI.human)
```

## Part II: Inference of cell-cell communication network
```{r inference of cell-cell com}
computeAveExpr(cellchat, features = c("Jag1","Notch2"), type =  "truncatedMean", trim = 0.01)

cellchat <- computeCommunProb(cellchat,
                              type =  "truncatedMean", trim = 0.01,
                              population.size = T)
# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
cellchat <- filterCommunication(cellchat, min.cells = 10)


# infer cell-cell comm at pathway level
cellchat <- computeCommunProbPathway(cellchat)

# aggregated cellcom network
cellchat <- aggregateNet(cellchat)


groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")


mat <- cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}


mat <- cellchat@net$count
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

## Part III: Visualization of cell-cell communication network
```{r comm network}
pathways.show <- c("NOTCH") 
# Hierarchy plot
# Here we define `vertex.receive` so that the left portion of the hierarchy plot shows signaling to fibroblast and the right portion shows signaling to immune cells 
vertex.receiver = seq(1,4) # a numeric vector. 
netVisual_aggregate(cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver)
# Circle plot
par(mfrow=c(1,1))
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "circle")

# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(cellchat, signaling = pathways.show, color.heatmap = "Reds")

netAnalysis_contribution(cellchat,
                         signaling = pathways.show)+
  theme(aspect.ratio = 1)

pairLR.NOTCH <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.NOTCH[13,] # show one ligand-receptor pair
# Hierarchy plot
vertex.receiver = seq(1,4) # a numeric vector
netVisual_individual(cellchat, signaling = pathways.show,  pairLR.use = LR.show, vertex.receiver = vertex.receiver)
#> [[1]]
# Circle plot
netVisual_individual(cellchat, signaling = pathways.show, pairLR.use = LR.show, layout = "circle")

netVisual_individual(cellchat, 
                     sources.use = "Myofibroblasts",
                     signaling = pathways.show,
                     thresh = 0.1,
                     pairLR.use = LR.show, layout = "circle")

saveRDS(cellchat, "objects/fibrotic_cellchat.rds")
```

